Graph Generators: State of the Art and Open Challenges

### 2019-12-18 main.tex
Irena's pass (plagiarism) pp 1-10
Typos fixed by Sherif

\acmformat{A. Bonifati, I. Holubov\'{a}, A. Prat-P\'erez, S. Sakr, 2019.
Modern Graph Generators: A Survey.}
% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

@@

\input{introduction}
%\input{preliminaries}
\input{comparison}
%\input{classification}
\input{generators}
%\input{comparison2}
%\input{data_sets}
\input{challenges}
\input{conclusion}

\begin{acks}
The authors would like to thank Dr. Kamesh Madduri for consultations and suggestions on the covered areas.
\end{acks}

>>>>

\acmformat{A. Bonifati, I. Holubov\'{a}, A. Prat-P\'erez, S. Sakr, 2019.
Graph Generators: State of the Art and Open Challenges.}
% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

@@

\section{Introduction}
\label{sec:intro}

%In recent years the term Big Data has become a phenomenon that breaks down borders of many technologies and approaches that have so far been acknowledged as mature and robust for any conceivable application. Gartner Inc.\footnote{\url{http://www.gartner.com/}} describes Big Data as ``\emph{high \textbf{v}olume, high \textbf{v}elocity, and/or high \textbf{v}ariety information assets that require new forms of processing to enable enhanced decision making, insight discovery and process optimization}''.

%\TODO{The abundance of interconnected data has fueled the design and implementation of graph generators reproducing real-world linking properties, or gauging the effectiveness of graph algorithms, techniques and applications manipulating these data.}

%One of the most popular data structures in the world of computer science are graphs.
Graphs are ubiquitous data structures that span a wide array of scientific disciplines
and are a subject of study in several subfields of computer science.
Nowadays, due to the dawn of numerous tools and applications manipulating
graphs, they are adopted as a rich data model in many data-intensive use
cases involving disparate domain knowledge, mathematical foundations and
computer science. Interconnected data is oftentimes used to encode domain-specific
use cases, such as recommendation networks, social networks,
protein-to-protein interactions, geolocation networks, and fraud detection analysis networks, to
name a few.
% ANGELA: I do not think we should cite Neo4J, which is a graph database.
%They enable to represent complex information in a simple and yet powerful way supported by a strong mathematical background. With the dawn of the new millennium which brought numerous novel technologies and applications, the graphs have become even more popular. There has appeared a number of use cases\footnote{\url{https://neo4j.com/use-cases/}} (such as fraud detection, recommendation engines, social networks etc.) where graphs represent the linked information.

In general, we can distinguish between two broad types of graph data sets: (1) a single large graph (possibly with several components), such as social networks  or Linked Data graphs, and (2) a large set of small graphs (e.g., chemical compounds\footnote{\url{https://pubchem.ncbi.nlm.nih.gov/}} or linguistics syntax trees\footnote{\url{https://catalog.ldc.upenn.edu/ldc99t42}}). Naturally, the algorithms used in these two classes differ a lot~\cite{DBLP:books/igi/Sakr2011}. In the former case we can search, e.g., for communities and their features or shortest paths, while in the latter case we usually query for supergraphs, subgraphs, or graphs similar to a given graph pattern. In both cases, as in the other fields, quite often the respective real-world graph data is not publicly available (or simply does not exist when a particular method for data manipulation is proposed).
Even in the cases in which real data is abundant, many algorithms and techniques need to be tested on various orders of magnitudes of the graph sizes thus leading to the inception of
configurable graph generators that reproduce real-world graph properties and provide unique tuning opportunities for
algorithms and tools handling such data.
%As a consequence, also in the world of graphs there has appeared a number of graph data generators which enable to  reproduce the respective real-world linking properties and gauge the effectiveness of a particular algorithm.

In this survey, we provide a detailed overview of the state-of-the-art
modern
graph generators by focusing on those that are pertinent and suitable for
data-intensive tasks. Our aim is to cover a wide range of currently popular
areas of graph data processing. In particular, we consider graph generation
in the areas of Semantic Web, graph databases, social networks, and community detection, along with general graphs. Despite the disparate requirements of modern graph generators throughout these communities, we analyze them under a common umbrella, reaching out the functionalities, the practical usage, and the supported operations of graph generators. The reasons for this scope and classification are as follows:

\begin{enumerate}
  \item Despite the differences of the covered areas, the requirements for
modern graph data generators can be similar in particular cases. Reusing or learning from tools in other fields can thus bring new opportunities for both researchers and practitioners.
  \item The selected classification is serving the need of providing scientists, researchers, and practitioners with the right data generator at hand for their work.
\end{enumerate}

To conclude the comparative study and provide a comprehensive view of the
field,  we also overview the most popular real-world data sets used in the respective covered areas and discuss open challenges and missing requirements of modern
graph generators in view of identifying their future extensions to new emerging fields.

\paragraph*{Contributions} Our survey revisits the representatives of
modern graph data generators and summarizes their main characteristics. The detailed review
and analysis make this paper useful for stimulating new graph data generation mechanisms as well as
serving as a main technical reference for selecting suitable solutions. In particular, the introductory
categorization and comparative study enables the reader to quickly get his/her bearings in the field and
identify a subset of generators of his/her interest. Since we do not limit the survey to a particular area of graph data management,  the reader can get a broader scope in the field. Hence, while practitioners can find a solution in another previously unexpected area, researchers can identify new target areas or exploit successful results in new fields. Last but not least, %to provide a global view of the state-of-the-art, we also provide an overview of most commonly used real-world testing graph datasets. And, finally,
we identify general open problems of graph data synthesis and indicate possible solutions to show that it still forms a challenging and promising research area.

\paragraph*{Differences with prior surveys}

To the best of our knowledge, this paper is the first to survey the broad
landscape of graph data generators spanning different data-intensive applications and
targeting many computer science subfields. In particular, we cover Semantic Web, graph databases, social networks, and community detection, along with general graphs. However, the literature is still lacking a comprehensive study of graph generators  for many of the specific subfields mentioned above.

A limited subset of graph database
generators, parallel and distributed graph processing generators, along
with a few of the Semantic Web data generators presented in our survey have
been discussed in a related book chapter
\cite{BFHI18} while cross-comparing them with respect to input, output,
supported workload, data model and query language, along with the
distinguished chokepoints. However, the provided classification is
inherently database-oriented.
In our work, we provide a more comprehensive and broader classification
 that serves the purpose of letting any
researcher or practitioner interested in data generation  to be able to make a better
choice of the desired graph
generator based on its functional and goal-driven features (such as the application domain,
the supported operations, and the key configuration options).
Moreover, in contrast with \cite{BFHI18}, our work encompasses graph generators of several
diverse communities, not limiting its scope to a few generators of the database
and graph processing communities.

Graph generators matching graph patterns used in data mining have been
studied in \cite{Chakrabarti:2006:GML:1132952.1132954},
focusing on mostly occurring patterns, such as power laws, size of graph diameters,
and community structure. The considered graph generators are compared in
terms of graph type, degree distributions, exponentiality, diameter, and
community effects. We refer the reader to this survey for taxonomies
involving these properties, whereas we provide here a functionality-driven
taxonomy across all the categories of graph generators that we consider.
We also point out that this survey is outdated as it does not consider the
generators that appeared in the last decade.
We fill the gap of more recent social network generators
in Section~\ref{sec:generators_socialnetworks}, as well as more recent representatives of the other categories.

%\TODO{Check with Arnau}

Aggarwal and Subbian~\cite{AggarwalS14} have surveyed the evolution of analysis in
graphs, by primarily focusing on data mining maintenance methods and on analytical
quantification and explanation of the changes of the underlying networks.
A brief discussion on evolutionary
network data generators is carried out in
the paper. The data generation of evolutionary networks is based on the
shrinking diameters and Densification Power Law (DPL), i.e., community-guided
attachment properties~\cite{Leskovec:2005:GOT:1081870.1081893}.
Generation of graphs tackling Kronecker
recursion with recursive tensor multiplication~\cite{AkogluMF08} is then considered in the above survey.
We refer the readers to the aforementioned survey for evolutionary network generators and further
discuss the open challenges of evolving graph data in Section~\ref{sec:challenges}.
%\TODO{Better highlight the differences when Sherif's part is in.}

%\TODO{Outreach other surveys in other communities.}

\paragraph*{Outline} The rest of the text is structured as follows: Section~\ref{sec:comparison}
provides the opening categorization and comparison of the generators. Section~\ref{sec:generators}
provides an overview of the existing graph data generators and their main features in the frame of
the proposed categories.
%To provide a complete overview,  we list the most commonly used testing data sets for the particular categories in Section~\ref{sec:data_sets}.
In Section~\ref{sec:challenges}, we highlight some of the challenges and open problems of graph
data synthesis before we conclude in Section~\ref{sec:conclusion}.

%\input{preliminaries}
\section{Classification and Comparative Study}
\label{sec:comparison}

In order to provide a general preview of the generators and  to enable finding the target solutions easily, we start the survey with a  classification and comparative study of the existing tools. In general, there are various ways to classify them. We first offer an overview of the approaches used in state-of-the-art work after which we then introduce our approach. As mentioned in the Introduction, since this survey is unique especially in terms of scope, our classification and comparative strategy differs as well.

The graph data generators can be classified using various distinct criteria. For example,~\cite{DBLP:conf/sdm/ChakrabartiZF04} introduces two categories -- degree-based and procedural generators. In general, \emph{degree-based generators} (e.g., Barabasi-Albert model~\cite{Barabasi99emergenceScaling}) are commonly attempting to find a graph that matches it, but without providing any information about the graph or attempting to match other graph features (e.g., eigenvalues, small diameter, etc). On the other hand, \emph{procedural generators} (e.g., R-MAT~\cite{DBLP:conf/sdm/ChakrabartiZF04}) are commonly targeting simple techniques to produce graphs that are matching the characteristics of the real-world graphs (such as, e.g., the power law degree distribution).

Paper~\cite{Chakrabarti:2006:GML:1132952.1132954} introduces five categories of graph models that can be synthesized: (1) \emph{random graph models} (e.g., Erd\"{o}s-R\'{e}nyi~\cite{Erdos:1960}) generated by a random process, (2) \emph{preferential attachment models} (e.g., Barabasi-Albert model) which try to model the power law from the preferential attachment viewpoint, (3) \emph{optimization-based models} (e.g., HOT model~\cite{PhysRevLett.84.2529}) resulting from the idea that power laws can result from tolerance to risks and resource optimizations, (4) \emph{tensor-based models} (e.g., R-MAT) targeting a trade-off between low number of model parameters and efficiency, and (5) \emph{internet-specific models} corresponding to hybrids using ideas from the other categories in order to suit the specific features of the graphs.

The type of the generator can also be influenced by the benchmark involving it, whereas we can distinguish, e.g., \emph{domain-specific} benchmarks, \emph{application-specific} benchmarks, \emph{workload-driven} benchmarks,  \emph{microbenchmarks} etc.

%\paragraph{Erd\"{o}s-R\'{e}nyi} One of the first and still very popular approaches with lots of variations and modifications is the random graph model proposed in~\cite{Erdos:1960}. Given a positive integer $n$ and a probability value $0 \leq p \leq 1$, the graph $G(n,p)$ is an undirected graph on $n$ vertices whose edges are chosen s.t. for all pairs of vertices $u,v$ there is an edge $(u,v)$ with probability $p$.
%
%\paragraph{Barabasi-Albert} The Barabasi-Albert (BA) model~\cite{Barabasi99emergenceScaling} proposes that structure emerges in network topologies as the result of two processes -- growth and preferential attachment. The BA model starts with a small set of nodes and grows the network as nodes and edges are added over time. Preferential attachment is based on the idea that the probability of connecting to a node is proportional to the current degree of that node (i.e., ``the rich get richer'').
%
%\paragraph{HOT} The highly optimized tolerance (HOT)~\cite{PhysRevLett.84.2529} model is introduced  using a ``forest fire'' example as follows: Suppose we have a forest which is prone to forest fires. Each portion of the forest has a different chance of starting the fire. We wish to minimize the damage by assigning (a limited amount of) resources such as firebreaks at different positions in the forest. More precisely, we have $n$ possible events, each with an associated probability $p_i, 1 \le i \le n$. Each event can lead to some loss $l_i$, which is a function of the resources $r_i$ allocated for that event $(l_i = f (r_i))$, whose total number is limited ($\sum_i r_i \le R$ for some given $R$). The aim is to minimize the expected cost $\sum_i p_i l_i$. The authors show that resource placement is related to the probability distribution $p_i$ by a power law and the probability of events which cause a loss greater than some value $k$ is related to $k$ by a power law too.
%
%
%\paragraph{R-MAT} The R-MAT (Recursive MATrix)~\cite{DBLP:conf/sdm/ChakrabartiZF04} model can be used to create graphs with community structure, power law degree distributions, and a small diameter. It generates weighted, directed and bipartite graphs. Graph generation is modeled with a matrix recursion -procedure in which the adjacency matrix (empty at the beginning) is recursively subdivided into four equal-sized partitions and edges are distributed across the partitions with unequal probabilities.

%In Figure~\ref{fig:classification} ...
%
%\begin{figure}
%\centering
%\includegraphics[width=0.8\textwidth]{classification.png}
%\caption{Classification of graph data generators}
%\label{fig:classification}
%\end{figure}

\subsection{Classification}

At first we classify the generators on the basis of the respective application domains or user communities. In particular we distinguish (1) general graphs, (2) Semantic Web, (3) graph databases, (4) social networks, and (5) community detection. The selected classes are not rigorously defined (e.g., they are not disjoint as we will show later), but they correspond to the currently most active research areas. Thus we believe that they form a natural first acquaintance for the reader.

In Table~\ref{tab:comparisonCharacteristics}   we overview the key characteristics of the data generators clustered according to the respective application domains.\footnote{``GDBs'' stands for graph databases, ``SNs'' stands for social networks, and ``Co'' stands for community detection. Value ``$-$'' means that the information is not available or relevant.} In particular, we show:

\begin{itemize}
\item Characteristics of the \textbf{domain}:
    \begin{itemize}
      \item its \textit{type} (column ``Type''), i.e.,  fixed, specified using a schema, or extracted from input data, and
      \item the particular \textit{target} domain, or, in case of a generic tool, the chosen sample domain (column ``Target/sample'').
\end{itemize}
\item Characteristics of \textit{read}/\textit{update} \textbf{operations} (columns ``Read'' and ``Update''), i.e., whether the set of operations is fixed/generated, if it involves operation mixes (i.e., sets/sequences of operations), or if templates of operations are supported.
\item Key \textbf{configuration} options:
  \begin{itemize}
    \item whether the generator deals only with structure, or also with \emph{properties} (column ``Pro.'') of the graph (Y/N feature),
    \item supported types of \emph{distributions} (column ``Distributions'') used for generating of the data,
    \item \textit{output format} (column ``Output'') of the produced graph, and
    \item  whether the generator is \textit{distributed} (column ``Dis.'') and thus enables more efficient data generation (Y/N feature).
  \end{itemize}
\end{itemize}

\begin{sidewaystable}
\scriptsize
\centering
\tbl{Key characteristics of the generators} {
\begin{tabular}{| c | l | p{2cm} |  p{2cm} | p{1.3cm} |  l | l | p{3.2cm} | p{2cm} | l | }
 \hline
           &   & \multicolumn{2}{c}{\textbf{Domain}}
               & \multicolumn{2}{|c|}{\textbf{Operations}}
               & \multicolumn{4}{c|}{\textbf{Configuration}}
               \\ \hline
           &  \textbf{Generator}
               & \textbf{Type}
               & \textbf{Target/sample}
               & \textbf{Read}
               & \textbf{Update}
               & \textbf{Pro.}
               & \textbf{Distributions}
			   & \textbf{Output}
               & \textbf{Dis.}
               \\ \hline
\hline   % ====================== general graphs
\multirow{7}{*}{\rot{\textbf{General}}}
  & Preferential attachment & -- & -- & N & N & Y & power-law & edge-list &  N  \\
\cline{2-10}
   & R-MAT & -- & -- & N & N & Y & power-law & edge-list &  N  \\
\cline{2-10}
  & HPC Scal. Graph Anal. & fixed & -- & fixed & N& N& uniform &edge-list & N   \\
\cline{2-10}
  & GraphGen & -- & -- & N & N & Y& user-defined  & node/edge-list & N   \\
\cline{2-10}
%  & Graph 500  & -- & -- & Y & N & Y & power-law & edge-list &  Y  \\
%\cline{2-10}
  & BTER & -- & -- &   N & N & N & user-defined & edge-list & Y  \\
\cline{2-10}
  & Darwini & -- & -- &   N & N & N & user-defined &  edge-list & Y   \\
\cline{2-10}
  & RTG & -- & -- &   N & N & N & power-law &  edge-list & N  \\
%  & Watts-Strogatz~\cite{watts1998collective} & fixed & & Y & N & Y &  & &  N  \\
%\cline{2-10}
\hline
\hline %  ======================  Semantic web
\multirow{19}{*}{\rot{\textbf{Semantic Web}}}
 & LUBM & fixed & university  & fixed & N & Y & random (LCG) &  RDF & N   \\
\cline{2-10}
 & LBBM & extracted & Lehigh university BibTeX  & N & N & Y & Monte Carlo &  RDF & N   \\
\cline{2-10}
 & UOBM & fixed & university  & fixed & N & Y & random &  RDF & N   \\
\cline{2-10}
 & IIMB & fixed & movies  & N & N & Y & random &  RDF & N   \\
\cline{2-10}
 & BSBM & fixed & e-commerce  & fixed & N & Y & mostly normal &  RDF, relational & N   \\
\cline{2-10}
 & SP$^2$Bench & fixed & DBLP  & fixed & N & Y & based on DBLP  & RDF & N   \\
\cline{2-10}
 & \cite{Duan:2011:AOC:1989323.1989340} & extracted & -- & N & N & Y & -- &  RDF & N    \\
\cline{2-10}
 & DBPSB & extracted & DBpedia &  templates & N & Y & random &  RDF & N   \\
\cline{2-10}
 & LODIB & fixed & e-commerce &  N & N & Y & 44 types &  RDF & N   \\
\cline{2-10}
 & Geographica & fixed & OpenStreetMap  & fixed + templates  & N & Y & -- &  RDF & N   \\
\cline{2-10}
 & WatDiv & schema-driven & user-defined  & templates & N & Y & uniform, normal, Zipfian &  RDF & N   \\
\cline{2-10}
 & RBench & extracted & DBLP, Yago  & templates & N & Y & from real-world data &  RDF & N  \\
\cline{2-10}
 & S2Gen & schema-driven & social network & Y & N & N & user-defined & RDF & N     \\
\cline{2-10}
 & RSPLab & schema-driven & agnostic & Y & Y & N & user-defined & RDF & N     \\
\cline{2-10}
 & LDBC SPB & fixed & media  & mixes & N & Y & power-law, skewed values, value correlation &  RDF & N  \\
\cline{2-10}
 & LinkGen & schema-driven & user-defined & templates & N  & N & Gaussian, Zipfian & RDF & N\\
\hline
\hline  % ======================  graph databases
\multirow{4}{*}{\rot{\textbf{GDBs}}}
  & XGDBench & fixed  & social network  & generated & generated & Y & power-law &  MAG &  Y  \\
\cline{2-10}
  & gMark & schema-driven &  user-defined  & generated &  N  & Y & uniform, normal, Zipfian &  N-triples & N    \\
\cline{2-10}
  & graphGen & pattern-driven & user-defined  & -- & -- & Y & -- &  GraphJson, CypherQueries & N   \\
\hline
\hline %  ======================   social networks
\multirow{8}{*}{\rot{\textbf{SNs}}}
 & \cite{Barrett:2009:GAL:1995456.1995598} & fixed & social network & N & N & Y & simulation-driven & impl. NA &  -- \\
\cline{2-10}
 & \cite{Yao2011} & fixed & social network & N & N & N & power-law & impl. NA & --  \\
\cline{2-10}
 & LinkBench & fixed & social network & generated & generated & Y & Facebook & impl. NA & -- \\
\cline{2-10}
 & S3G2 & fixed & social network  & N & N  & Y & Facebook  & CSV, RDF & Y   \\
\cline{2-10}
 & SIB & fixed & social network &  mixes & mix & Y & from real-world data &  RDF & N   \\
\cline{2-10}
 & \cite{Sukthankar-SocialInfo2014} & schema-driven & social-network & N & N & Y & power-law & CSV & N   \\
\cline{2-10}
 & LDBC SNB  & fixed & social network &  generated & generated  & Y & Facebook &  CSV, RDF & Y     \\
\cline{2-10}
  & \cite{Nettleton2016} & schema-driven & social network & N & N & Y & power-law & impl. NA & --  \\
%\cline{2-10}
% & SNG  & Y & Y &   N &  & Y & power-law & ? &  ?   \\
%\cline{2-10}
% & Forest Fire  & N & Y   & N &  & N & power-law &  ? & N   \\
%\hline
%\hline   % ======================   graph analytics
%\multirow{1}{*}{\rot{\textbf{A}}}
%  & HPC Scal. Graph Anal. & fixed & -- & fixed & N& N& uniform &edge-list & N   \\
%\cline{2-10}
%  & Graphalytics & extracted & social network & fixed& N& N & power-law & CSV, RDF & N   \\
\hline
\hline   %  ======================  community detection
\multirow{4}{*}{\rot{\textbf{Co}}}
  & \cite{danon2005comparing} & -- & -- & Y & N & N & uniform & edge-list &  N    \\
\cline{2-10}
  & LFR & -- & -- & Y & N & N & power-law & edge-list  & N      \\
\cline{2-10}
  & LFR-Overlapping & -- & -- & Y & N & N & power-law & edge-list & N     \\
\cline{2-10}
  & Stochastic Block Models & -- & -- & Y & N & N & user-defined & edge-list & Y     \\
\hline
%\hline   %  ======================  streaming
%\multirow{2}{*}{\rot{\textbf{St}}}
%  & S2Gen & schema-driven & social network & Y & N & N & user-defined & RDF & N     \\
%\cline{2-10}
%  & RSPLab & schema-driven & agnostic & Y & Y & N & user-defined & RDF & N     \\
%\cline{2-10}
%  & & & & & & & &  &   \\
\end{tabular} }
\label{tab:comparisonCharacteristics}
\end{sidewaystable}

\paragraph{Number of Generators, Size and Output Format of the Data}
As we can see, the biggest set of available generators can be found in the Semantic Web application domain, probably due to its recent popularity and a number of research groups dealing with this topic. None of these generators is natively implemented in a distributed manner and thus primarily generating  output at the Big Data scale. On the other hand, the \emph{Linked Open Data} (LOD) is expected to be large in general; however, this is the case of the whole \emph{LOD Cloud}\footnote{\url{https://lod-cloud.net/}}, but not necessarily of the particular data sets forming it.

The second large group of generators also corresponds to a popular application domain -- social networks. In this case the size of the output graph is important if we require realistic features of the result. However, surprisingly many of the proposals do not provide an implementation at all and amongst the others there is a high percentage of those that are not highly scalable.

In case of the other application-specific domains the amount of generators is relatively small. As we will show in Section~\ref{sec:overlapping}, the situation is not so critical. Some of the application-specific generators can be re-used also in other application domains or the general graph generators can be used.

Considering the output format of the data, as expected, the generators produce data in a standard format (e.g., RDF) or in a reasonable graph-related form (e.g., edge list).

\paragraph{Domain}

If we consider the features of the particular domain of the generators, in most cases it is expectably \emph{fixed}, i.e., it is pre-defined (describing, e.g., a social network) and cannot be modified. It is the simplest option, but it does mean that the generator is simple too -- it can still focus on other complex aspects of the output. While the \emph{schema-driven} approaches enable to influence the target domain using a user-supplied schema, there also exist approaches where the domain is \emph{extracted} from sample data. The particular target (in case of fixed) or sample (in case of schema-driven or extracted) domains do not provide very rich areas. They either correspond to the respective application domains (like in the case of social networks), or they are based on well-known and commonly used data sets (such as, e.g., DBLP~\cite{dblp} or DBpedia~\cite{Bizer:2009:DCP:1640541.1640848}). Except for general graphs without any domain, in all other cases we can find a flexible representative with schema-driven/pattern-driven domain or a domain extracted from sample data.

\paragraph{Operations}
In the case of operations, most commonly the respective generators are accompanied with a set of fixed read operations or sequences of operations (query mixes) representing typical behavior of a user. In some cases this aspect is more flexible -- either query templates are used or the operations are generated, e.g., in order to access most of the data in the generated graph. On the other hand, update operations are provided only in a small amount of cases. As we will discuss in Section~\ref{sec:evolving}, the more general problem of evolving graphs is still an open issue.

\paragraph{Configuration}
A natural feature of the generators is to provide as realistic graphs as possible. Hence, most of them focus not only on the structure of the output graph, but also the properties. For the purpose of generating graphs with near real-world characteristics various distributions are used, such as power-law, Zipfian etc. Especially interesting are distributions extracted from real-wold data (such as, e.g., Facebook in case of the social network domain).

\subsection{Overlapping}
\label{sec:overlapping}

As we have mentioned, the basic classification of the generators that we have used in this paper is relatively vaguely based on the current application domains or research areas. In addition, some of the generators are either general, and thus can be used universally, or have features applicable in more than one domain/research area. So the classes we use can overlap, as depicted in Table~\ref{tab:overlapping}.

\begin{table}
\scriptsize
\centering
\tbl{Overlapping of classes of generators} {
\begin{tabular}{| c | l | l | l | l | l | l | l | }
 \hline
           &  \textbf{Generator}
               & \textbf{\rot{General}}
               & \textbf{\rot{Semantic Web}}
               & \textbf{\rot{Graph databases\ }}
               & \textbf{\rot{Social networks}}
               %& \textbf{\rot{Analytics}}
               & \textbf{\rot{Community detection\ }}
               %& \textbf{\rot{Streaming}}
               \\ \hline
\hline   % ====================== general graphs
\multirow{6}{*}{\rot{\textbf{General}}}
  & Preferential attachment & {\bf x} & & &  &  \\
\cline{2-7}
   & R-MAT    & {\bf x} & & &  &  \\
\cline{2-7}
   & HPC Scal. Graph Anal.  & {\bf x} & & &  &  \\
\cline{2-7}
  & GraphGen  & {\bf x} & & x &  & \\
\cline{2-7}
%  & Graph 500 & {\bf x} & & &  & \\
%\cline{2-7}
  & BTER      & {\bf x} & & &  & \\
\cline{2-7}
  & Darwini   & {\bf x} & & &  & \\
\cline{2-7}
  & RTG   & {\bf x} & & &  & x \\
\hline
\hline %  ======================  Semantic web
\multirow{16}{*}{\rot{\textbf{Semantic web}}}
 & LUBM  & & {\bf x} & &  &  \\
\cline{2-7}
 & LBBM  & & {\bf x} & &  &  \\
\cline{2-7}
 & UOBM  & & {\bf x} & &  &  \\
\cline{2-7}
 & IIMB & & {\bf x} & &  & \\
\cline{2-7}
 & BSBM & & {\bf x} & &  &  \\
\cline{2-7}
 & SP$^2$Bench & & {\bf x} & &  &  \\
\cline{2-7}
 & \cite{Duan:2011:AOC:1989323.1989340} & & {\bf x} & &  &  \\
\cline{2-7}
 & DBPSB & & {\bf x} & &  &  \\
\cline{2-7}
 & LODIB & & {\bf x} & &  &  \\
\cline{2-7}
% & SIB & & {\bf x} & &  &  \\
%\cline{2-7}
 & Geographica & & {\bf x} & &  &  \\
\cline{2-7}
 & WatDiv & & {\bf x} & &  &  \\
\cline{2-7}
 & RBench & & {\bf x} & &  &  \\
\cline{2-7}
  & S2Gen & & {\bf x} & &  &  \\
\cline{2-7}
  & RSPLab & & {\bf x} & &  &  \\
\cline{2-7}
 & LDBC SPB & & {\bf x} & &  &  \\
\cline{2-7}
 & LinkGen & & {\bf x} & &  &  \\
\hline
\hline  % ======================  graph databases
\multirow{3}{*}{\rot{\textbf{GDBs}}}
  & XGDBench & & & {\bf x} & x  &  \\
\cline{2-7}
  & gMark & & x & {\bf x} & x  &  \\
\cline{2-7}
  & graphGen & & & {\bf x} &  &  \\
\hline
\hline %  ======================   social networks
\multirow{8}{*}{\rot{\textbf{SNs}}}
 & \cite{Barrett:2009:GAL:1995456.1995598} & & & & {\bf x}  & \\
\cline{2-7}
 & \cite{Yao2011}  & & & & {\bf x}  & \\
\cline{2-7}
 & LinkBench  & & & x & {\bf x}  & \\
\cline{2-7}
 & S3G2  & & x & x & {\bf x}  & \\
\cline{2-7}
 & SIB & & x & & {\bf x}  &  \\
\cline{2-7}
 & \cite{Sukthankar-SocialInfo2014}  & & & & {\bf x}  & \\
\cline{2-7}
 & LDBC SNB   & & x & x & {\bf x}  & \\
\cline{2-7}
  & \cite{Nettleton2016}  & & & & {\bf x}  &  \\
%\hline
%\hline   % ======================   graph analytics
%\multirow{1}{*}{\rot{\textbf{A}}}
%  & HPC Scal. Graph Anal.  & & & & &  \\
%\cline{2-7}
%  & Graphalytics & & & & &  \\
\hline
\hline   %  ======================  community detection
\multirow{4}{*}{\rot{\textbf{Co}}}
  & \cite{danon2005comparing} & x & & & &  {\bf x} \\
\cline{2-7}
  & LFR & x & & & &  {\bf x}  \\
\cline{2-7}
  & LFR-Overlapping & x & & & &  {\bf x} \\
\cline{2-7}
  & Stochastic Block Models & x & & & &  {\bf x} \\
%\cline{2-7}
%  & & & & &  \\
\hline
%\hline   %  ======================  streaming
%\multirow{2}{*}{\rot{\textbf{St}}}
%  & S2Gen & & & & & & & {\bf x}\\
%\cline{2-7}
%  & RSPLab & & & & & & & {\bf x} \\
%%\cline{2-7}
%%  & & & & & & \\
%\hline
\end{tabular} }
\label{tab:overlapping}
\end{table}

For example, many general or domain-agnostic graph generators, such as the
preferential attachment~\cite{Barabasi99emergenceScaling} or R-MAT, are typically used to test graph analytics frameworks when large real graphs are not available. %Actually, Graphalytics uses R-MAT as one as its graph generators for benchmarking graph analytics.

Similarly, some social network graph generators such as LDBC SNB, S3G2 or
LinkBench, can be used to test graph databases. In the case of the first two,
even though they are designed not to be specific to any type of technology,
the graph databases are their main target.  Additionally, they also provide serializers for
RDF, thus they can also be used to test RDF systems.

In the case of LinkBench, nothing prevents the user to load the generated graph
in a (graph) database (e.g., Facebook uses MySQL in~\cite{Armstrong:2013:LDB:2463676.2465296}) to
test a workload similar to Facebook and extend and complement it with more
graph queries like those in LDBC SNB.

Generators for community detection aim, in general, at creating graphs with a
more realistic structure (graphs with communities of nodes where the density of
edges is larger internally than externally). Even though these generators are,
in general, used to test community detection algorithms (they generate also the
expected communities in the graph), some studies also use them for general graph
purposes or to test graph analytics algorithms besides community detection.

%\input{classification}
\section{Graph Data Generators}
\label{sec:generators}

In this section, we discuss the various graph data generators based on the classification introduced before in more detail. For each category, we first describe the key features of each of the representative examples and summarize their strengths and weaknesses. The goal is to offer a detailed information about each of the tools in the context of its competitors from the same domain.

\subsection{General Graphs}
\label{sec:generators_general}

We start by focusing on approaches that have been designed for dealing with the
generation of general graph data that is not aimed at a particular application
domain. In general, such generators focus on reproducing properties observed in
real graphs regardless of their domain such as the degree distribution, the
diameter, the presence of a large connected component, a community structure or
a significantly large clustering coefficient.

%Currently, there exists a number of tools which involve a kind of
%general graph data generator, such as gtools from projects nauty and
%Traces~\cite{gtools} or the Stanford GraphBase~\cite{GraphBase}. We, however,
%focus on primarily generating/benchmarking projects targeting the Big Data
%world.

\paragraph {Preferential Attachment} Barabasi and
Albert~\cite{Barabasi99emergenceScaling} introduced a graph generation model
that relies on two main mechanisms. The first mechanism is continuously
expanding the graphs by adding new vertices. The second mechanism is to
preferentially attach the new vertices   to the nodes/regions that are already
well connected. So, in this approach, the generation of large graphs is governed
by standard, robust self-organizing mechanisms that go beyond the
characteristics of individual applications.

\paragraph {R-MAT} R-MAT (\emph{R}ecursive \emph{Mat}rix) is a procedural
synthetic graph generator which is designed to generate power-law degree
distributions~\cite{DBLP:conf/sdm/ChakrabartiZF04}.  The generator is recursive
and employs a fairly small number of parameters.  In principle, the strategy of
this generator is to achieve simple means to produce graphs whose properties correspond to
properties of the real-world graphs. In particular, the design goal of R-MAT is to
produce graphs which mimics the degree distributions, imitate a community
structure and have a small diameter. R-MAT can generate weighted, directed and
bipartite graphs.

\paragraph{HPC Scalable Graph Analysis Benchmark} The HPC Scalable Graph
Analysis Benchmark~\cite{HPCgraph,Bader:2005:DIH:2099301.2099360} consists of a
weighted, directed graph that has a power-law distribution and four related
analysis techniques (namely graph construction, graph extraction with BFS,
                     classification of large vertex sets, and graph analysis
                     with betweenness centrality). The generator has the
following parameters: the number of nodes, the number of edges, and maximum weight of
an edge. It outputs a list of tuples containing identifiers of vertices of an
edge (with the direction from the first one to the second one) and weights
(positive integers  with a uniform random distribution) assigned to the edges of
the multigraph.  The algorithm of the generator is based on R-MAT.
%To reduce the data
%locality, the vertex ids are randomly permuted and the output
%tuples are shuffled. A related project from the same authors
%developed for the 9th DIMACS Shortest Paths Challenge is GTgraph~\cite{GTgraph}.
%t involves three types of graphs: (1) Erd\"{o}s-R\'{e}nyi random graphs, (2)
%different graphs from the DARPA HPCS SSCA\#2 graph theory benchmark
%(version 1.0), and (3) small-world graphs generated using the R-MAT graph
%generator.

\paragraph{GraphGen} For the purpose of testing the scalability of an indexing
technique called FG-index~\cite{Cheng:2007:FTV:1247480.1247574} on the size of the
database of graphs, their average size and average density, the authors have
also implemented a synthetic generator called
GraphGen\footnote{https://www.cse.ust.hk/graphgen/}. It relies on data generation code for associations and sequential
patterns provided by IBM\footnote{From 1996, no longer available at
\url{http://www.almaden.ibm.com/cs/projects/iis/hdb/Projects/data
mining/mining.shtml}}. GraphGen yields a collection of undirected, labeled and
connected graphs. It addresses the performance evaluation of frequent subgraph mining
algorithms and graph query processing algorithms. The result is represented as a
list of graphs, each consisting of a list of nodes along with a list of edges.
